### 2011-01-11 doc/man/rt1d_manual.tex
Added bandpass averaging discretization technique in RadiationSource.

%%% rt1d

\documentclass[letterpaper,titlepage,12pt]{article}

%%% Preamble

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\marginparsep}{0in}
\setlength{\marginparwidth}{0in}

\usepackage{amsmath, amsthm, amssymb}
\numberwithin{equation}{section}
\usepackage{graphicx}
\usepackage{pslatex}

%%% Beginning of Document

\begin{document}

\author{Jordan Mirocha}

\title{\Large {\bf rt1d: User's Manual}}
\date{Last Updated: \today}
\maketitle

\setcounter{tocdepth}{2}
\tableofcontents
\newpage

%%% 
%% Methods
%%%
\section{Methods}

%%% 
%% Classes
%%%
\section{Classes}
In order of appearance in \texttt{rt1d.py}.

\subsection{\texttt{InitializeGrid}}
\subsection{\texttt{InitializeIntegralTables}}
\subsection{\texttt{RadiationSource}}
\subsection{\texttt{Radiate}}

%%% 
%% Start to Finish
%%%
\section{Start to Finish}

%%% 
%% Parameters
%%%
\section{Parameters}

\subsection{Initial Conditions}
\begin{description}

\item [\texttt{DensityProfile}]

\item [TemperatureProfile]

\end{description}

\subsection{Control Parameters}

\subsection{Source Parameters}

\end{document}