### 2012-05-23 doc/manual/rt1d_manual.tex
Minor changes. About to rework lookup table initialization and interpolation to accomodate higher (4-5) dimensional tables.

%%% rt1d

\documentclass[letterpaper,titlepage,12pt]{article}

%%% Preamble

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\marginparsep}{0in}
\setlength{\marginparwidth}{0in}

\usepackage{amsmath, amsthm, amssymb}
\numberwithin{equation}{section}
\usepackage{graphicx}
\usepackage{pslatex}
\usepackage{epsfig}
\usepackage{natbib}
\usepackage[pdfborder={0 0 1}]{hyperref}
\usepackage[htt]{hyphenat}

%%% START custom commands
\newcommand{\HI}{\text{HI}}
\newcommand{\HII}{\text{HII}}
\newcommand{\HeI}{\text{HeI}}
\newcommand{\HeII}{\text{HeII}}
\newcommand{\HeIII}{\text{HeIII}}
\newcommand{\nH}{n_{\text{H}}} 
\newcommand{\nHI}{n_{\text{HI}}} 
\newcommand{\nHII}{n_{\text{HII}}}  
\newcommand{\nHe}{n_{\text{He}}}
\newcommand{\xHI}{x_{\text{HI}}} 
\newcommand{\xHII}{x_{\text{HII}}} 
\newcommand{\xHeI}{x_{\text{HeI}}} 
\newcommand{\xHeII}{x_{\text{HeII}}} 
\newcommand{\xHeIII}{x_{\text{HeIII}}}  
\newcommand{\nHeI}{n_{\text{HeI}}}
\newcommand{\nHeII}{n_{\text{HeII}}}
\newcommand{\nHeIII}{n_{\text{HeIII}}}  
\newcommand{\nel}{n_{\text{e}}}  
\newcommand{\ntot}{n_{\text{tot}}}
\newcommand{\nnu}{$n_{\nu}$}
\newcommand{\ncol}{N_i}
\newcommand{\gHI}{\Gamma_{\text{HI}}}  
\newcommand{\gHeI}{\Gamma_{\text{HeI}}}
\newcommand{\gHeII}{\Gamma_{\text{HeII}}}
\newcommand{\aHII}{\alpha_{\text{HII}}}  
\newcommand{\aHeII}{\alpha_{\text{HeII}}}  
\newcommand{\aHeIII}{\alpha_{\text{HeIII}}}
\newcommand{\bHI}{\beta_{\text{HI}}} 
\newcommand{\bHeI}{\beta_{\text{HeI}}}  
\newcommand{\bHeII}{\beta_{\text{HeII}}}  
\newcommand{\xiHeII}{\xi_{\text{HeII}}}
\newcommand{\kB}{k_{\text{B}}}
\newcommand{\fheat}{f^{\text{heat}}}
\newcommand{\fheatE}{f_{(\nu - \nu_i)}^{\text{heat}}}
\newcommand{\fion}{f_i^{\text{ion}}}
\newcommand{\fionE}{f_{ij,(\nu - \nu_j)}^{\text{ion}}}
\newcommand{\Lbol}{\mathcal{L}_{\text{bol}}}
\newcommand{\spec}{\mathcal{N}}
\newcommand{\Heat}{\mathcal{H}}
\newcommand{\trec}{$t_{\text{rec}}$}
\newcommand{\Lbox}{L_{\mathrm{box}}}
\newcommand{\dx}{\Delta x}
\newcommand{\dd}{\text{d}}
\newcommand{\Htwo}{\mathrm{H}_2}
\newcommand{\drIF}{$\Delta r_{\mathrm{IF}}$}
\newcommand{\dTb}{$\delta T_b$}
\newcommand{\Nvec}{\mathbf{N}}
\newcommand{\sh}{\mathrm{sh}}
%%% END custom commands

%%% Beginning of Document
\begin{document}
\author{Jordan Mirocha}	
\title{\Large {\bf rt1d: Documentation}}
\date{Last Updated: \today}
\maketitle

% Table of contents
\setcounter{tocdepth}{2}
\tableofcontents
\newpage

%%%
%% Introduction
%%%
\section{Introduction}
Let's see what happens to the gas around stars and black holes, shall we? The
primary reference for this code is
\cite[\href{http://arxiv.org/abs/1204.1944}{arxiv:1204.1944}]{Mirocha2012} --
much of the text has been extracted from there.

%%% 
%% Methods
%%%
\section{Methods}

% Rate Equations
\subsection{The Rate Equations}
In general, the chemical and thermal evolution of gas surrounding a radiation
source is governed by a set of differential equations describing the number
densities of all ions and the temperature of the gas. Assuming a medium
consisting of hydrogen and helium only, we first solve for the abundances of
each ion via
\begin{align}
    \frac{d \nHII}{dt} & = (\gHI + \gamma_{\HI} + \bHI \nel) \nHI - \aHII \nel \nHII   \label{eq:HIIRateEquation} \\ 
    \frac{d \nHeII}{dt} & = (\gHeI + \gamma_{\HeI} + \bHeI \nel) \nHeI \nonumber + \aHeIII \nel \nHeIII \\  & - (\bHeII + \aHeII + \xiHeII) \nel \nHeII \label{eq:HeIIRateEquation} \\ 
    \frac{d \nHeIII}{dt} & = (\gHeII + \gamma_{\HeII} + \bHeII \nel) \nHeII  - \aHeIII \nel \nHeIII . \label{eq:HeIIIRateEquation}
\end{align}
Each of these equations represents the balance between ionizations of species
$\HI$, $\HeI$, and $\HeII$, and recombinations of $\HII$, $\HeII$, and
$\HeIII$. Associating the index $i$ with absorbing species, $i = \HI, \HeI,
\HeII$, and the index $i^{\prime}$ with ions, $i^{\prime} = \HII, \HeII,
\HeIII$, we define $\Gamma_i$ as the photo-ionization rate coefficient,
$\gamma_i$ as the secondary ionization rate coefficient, $\alpha_{i^{\prime}}$
($\xi_{i^{\prime}}$) as the case-B (dielectric) recombination rate
coefficients, $\beta_i$ as the collisional ionization rate coefficients, and
$\nel = \nHII + \nHeII + 2\nHeIII$ as the number density of electrons.

At each time-step, we also solve for the temperature evolution, $dT_k/dt$,
which
is given by
\begin{align}
    \frac{3}{2}\frac{d}{dt}\left(\frac{\kB T_k \ntot}{\mu}\right) & = \fheat  \sum_i n_i \Heat_i - \sum_i \zeta_i n_e n_i - \sum_{i^{\prime}} \eta_{i^{\prime}} n_e n_{i^{\prime}} \nonumber \\ & - \sum_i \psi_i n_e n_i - \omega_{\HeII} n_e \nHeII \label{eq:TemperatureEvolution} 
\end{align}
where $\Heat_i$ is the photo-electric heating rate coefficient (due to
electrons previously bound to species $i$), $\omega_{\HeII}$ is the dielectric
recombination cooling coefficient, and $\zeta_i$, $\eta_{i^{\prime}}$, and
$\psi_i$ are the collisional ionization, recombination, and collisional
excitation cooling coefficients, respectively. The constants in Equation
\ref{eq:TemperatureEvolution} are the total number density of baryons, $\ntot
= n_\mathrm{H} + n_{\mathrm{He}} + n_e$, the mean molecular weight, $\mu$,
Boltzmann's constant, $\kB$, and the fraction of secondary electron energy
deposited as heat, $\fheat$. We use the formulae in Appendix B of
\citet{Fukugita1994} to compute the values of $\alpha_i$, $\beta_i$, $\xi_i$,
$\zeta_i$, $\eta_{i^{\prime}}$, $\psi_i$, and $\omega_{\HeII}$.

% Ionization and Heating Rates
\subsection{Computing the Ionization and Heating Rates}
The most critical aspect of propagating the radiation field in our 1D
simulations is computing the ionization ($\Gamma_i$, $\gamma_{i}$) and heating
($\Heat_i$) rate coefficients accurately. In order to directly relate our
results to fully 3-dimensional radiative transfer calculations we have chosen
to adopt a photon-conserving (PC) algorithm nearly identical to those employed
by several widely used codes, like \textit{C$^2$Ray} \citep{Mellema2006}, and
\textit{Enzo} \citep{Wise2011}. Our code is able to compute $\Gamma_i$,
$\gamma_{i}$, and $\Heat_i$ in a non-photon-conserving (NPC) fashion as well,
to enable comparison with previous 1D work such as \citet{Thomas2008}. The two
formalisms are equivalent in the limit of very optically thin cells, a
condition that can be met easily in 1D calculations but is rarely
computationally feasible in 3D. For NPC methods, if the optical depth of an
individual cell is substantial, the number of ionizations in that cell will
\textit{not} equal the number of photons absorbed for that cell, i.e. photon
number will not be conserved. This problem was remedied by \citet{Abel1999},
who inferred the number of photo-ionizations of species $i$ in a cell from the
radiation incident upon it and its optical depth,
\begin{equation}
    \Delta \tau_{i,\nu} = n_i \sigma_{i,\nu} \Delta r .
\end{equation}    
It is most straightforward to imagine our 1D grid as a collection of
concentric spherical shells, each having thickness $\Delta r$ and volume
$V_{\sh}(r) = 4 \pi [(r + \Delta r)^3 - r^3] / 3$, where $r$ is the distance
between the origin and the inner interface if each shell. The ionization and
heating rates can then be related to the number of absorptions in any given
shell (thus preserving photon number), as
\begin{align}
    \Gamma_i & = A_i \int_{\nu_i}^{\infty} I_{\nu} e^{-\tau_{\nu}} \left(1 - e^{-\Delta \tau_{i,\nu}}\right) \frac{d\nu}{h\nu} \label{eq:PhotoIonizationRate} \\
    \gamma_{ij} & = A_j \int_{\nu_j}^{\infty} \left(\frac{\nu - \nu_j}{\nu_i}\right) I_{\nu} e^{-\tau_{\nu}} \left(1 - e^{-\Delta \tau_{j,\nu}}\right) \frac{d\nu}{h\nu} \label{eq:SecondaryIonizationRate} \\
    \Heat_i & = A_i \int_{\nu_i}^{\infty} (\nu - \nu_i) I_{\nu} e^{-\tau_{\nu}} \left(1 - e^{-\Delta \tau_{i,\nu}}\right) \frac{d\nu}{\nu} , \label{eq:HeatingRate}
\end{align}    
where we have defined the normalization constant $A_i \equiv
L_{\mathrm{bol}}/n_i V_{\sh}(r)$, and denote the ionization threshold energy
for species $i$ as $h\nu_i$.  $I_{\nu}$ represents the SED of radiation sources, and satisfies 
\begin{equation}
\int_{\nu} I_{\nu} d\nu = 1,
\end{equation}
such that $A_i I_{\nu}$ is the number of photons of frequency $\nu$ per atom (of species $i$) in a shell at radius $r$.

Equation \ref{eq:SecondaryIonizationRate} represents ionizations of species
$i$ due to fast secondary electrons from photoionizations of species $j$,
which has number density $n_j$, and ionization threshold energy, $h\nu_j$.
$\fion$ is the fraction of photo-electron energy deposited as ionizations of
species $i$. In the remaining sections we only include the effects of
secondary electrons when considering X-ray sources, which emit photons in the
range $10^2\mathrm{eV} < E < 10^4\mathrm{eV}$. In this regime, the values of
$\fheat$ and $\fion$ computed via the formulae of \citet{Shull1985} are
sufficiently accurate, but for radiation at lower energies where $\fheat$ and
$\fion$ have a stronger energy dependence, the lookup tables of
\citet{Furlanetto2010} would be more appropriate. The total secondary
ionization rate for a given species, $\gamma_i$, is the sum of ionizations due
to the secondary electrons from all species, $\gamma_i = \fion \sum_j
\gamma_{ij} n_j / n_i$.

The optical depth, $\tau_{\nu} = \tau_{\nu}(r)$, in the above equations is
the total optical depth at frequency $\nu$ due to all absorbing species, i.e.
\begin{align}
    \tau_{\nu}(r) & = \sum_i \int_0^r \sigma_{i,\nu} n_i(r^{\prime}) dr^{\prime} \nonumber \\
               & = \sum_i \sigma_{i,\nu} \ncol(r) \label{eq:OpticalDepth}
\end{align}
where $\ncol$ is the column density of species $i$ at distance $r$ from the
source. We calculate the bound-free absorption cross-sections using the fits
of \citet{Verner1996} throughout.

The values of $\Gamma_i$, $\gamma_i$, and $\Heat_i$ are completely
predetermined for a given radiation source, and as a result, can be tabulated
as a function of column density to avoid evaluating the
integrals in these expressions numerically `on-the-fly' as a simulation runs
\citep[e.g.][]{Mellema2006,Thomas2008}. Isolating the frequency dependent
components of Equations \ref{eq:PhotoIonizationRate}-\ref{eq:HeatingRate}, we
can define the integrals
\begin{align}
    \Phi_i (\tau_{\nu}) & \equiv \int_{\nu_i}^{\infty} I_{\nu} e^{-\tau_{\nu}} \frac{d\nu}{h\nu} \label{eq:PHI} \\
    \Psi_i (\tau_{\nu}) & \equiv \int_{\nu_i}^{\infty} I_{\nu} e^{-\tau_{\nu}} d\nu \label{eq:PSI},
\end{align}
allowing us to re-express the rate coefficients as
\begin{align}
    \Gamma_i & = A_i \left[\Phi_i(\tau_{\nu}) - \Phi_i(\tau_{i,\nu}^{\prime}) \right] \label{eq:Gamma_PhiPsi} \\
    \gamma_{ij} & = \frac{A_j}{h\nu_i} \left\{\Psi_j(\tau_{\nu}) - \Psi_j(\tau_{j,\nu}^{\prime}) - h \nu_j \left[\Phi_j(\tau_{\nu}) - \Phi_j(\tau_{j,\nu}^{\prime}) \right] \right\}  \label{eq:gamma_PhiPsi} \\
    \Heat_i & = A_i \left\{\Psi_i(\tau_{\nu}) - \Psi_i(\tau_{i,\nu}^{\prime}) - h\nu_i \left[\Phi_i(\tau_{\nu}) - \Phi_i(\tau_{i,\nu}^{\prime}) \right] \right\} \label{eq:Heat_PhiPsi},  
\end{align}
where $\tau_{i,\nu}^{\prime} \equiv \tau_{\nu} + \Delta \tau_{i, \nu}$. Later
references to ``continuous SEDs'' signify use of this technique, where the
integral values $\Phi_i$ and $\Psi_i$ are computed over a column density
interval of interest a priori using a Gaussian quadrature technique, rather
than on-the-fly via discrete summation.

Equations \ref{eq:Gamma_PhiPsi}-\ref{eq:Heat_PhiPsi} are completely general
for photon-conserving algorithms, whether the source SEDs are discrete or
continuous -- the only difference being for discrete SEDs, the integrals in
Equations \ref{eq:PHI}-\ref{eq:PSI} become sums over the number of frequencies
used, $n_{\nu}$. In practice, computing $\Gamma_i$, $\gamma_i$, and $\Heat_i$
is more straightforward than this for sources with discrete SEDs, as we can
simply count the number of ionizations caused by each individual frequency
group, and convert this into the amount of excess electron kinetic energy
available for further heating and ionization. When testing the accuracy of
discrete solutions in later sections we employ this method, where radiation is
emitted at $n_{\nu}$ frequencies, with each frequency $\nu_n$ carrying a
fraction $I_n$ of the bolometric luminosity. The photoionization and heating
coefficients for each individual frequency group can then be expressed as
\begin{align}
    \Gamma_{i,n} & = \frac{A_i I_n}{h \nu_n} e^{-\tau_{\nu_n}}(1 - e^{-\Delta \tau_{i,\nu_n}}) \label{eq:Gamma_simple}\\
    \gamma_{ij,n} & = \Gamma_{j,\nu_n} (\nu_n-\nu_j) / \nu_i \label{eq:gamma_simple} \\
    \Heat_{i,n} & = \Gamma_{i,\nu_n} h(\nu_n - \nu_i) \label{eq:Heat_simple}.
\end{align}    
The total rate coefficients can be found by summing each of these expressions
over all frequencies, $n=1,2,3...n_{\nu}$. These equations are identical to
Equations \ref{eq:Gamma_PhiPsi}-\ref{eq:Heat_PhiPsi} for the discrete SED
case, but are perhaps more intuitive.

% Secondary Electrons
\subsection{Additional Complexity: Energy Dependent Secondary Electron Treatment}
In the previous section (and in all of \citet{Mirocha2012}), we considered the asymptotic limit of \citet{Shull1985}, in which the fractional energy deposition of secondary electrons as heat, ionization, and excitation depends only on the hydrogen ionized fraction.  \citet{Ricotti2002} and \citet{Furlanetto2010} updated this work, providing (respectively) empirical fits and lookup tables of deposition fraction as a function of both ionized fraction and electron energy. These effects are most important at $E \lesssim 10^2$ eV.

To accommodate this model, we must redefine the quantities $\Phi_i$ and $\Psi_i$.  We will use a tilde to denote the new quantities related to secondary ionization, 
\begin{align}
    \widetilde{\Phi}_{ij} & \equiv \int_{\nu_j}^{\infty} \fionE I_{\nu} e^{-\tau_{\nu}} \frac{d\nu}{h\nu} \\
    \widetilde{\Psi}_{ij} & \equiv \int_{\nu_j}^{\infty} \fionE I_{\nu} e^{-\tau_{\nu}} d\nu ,
\end{align}
and a hat to denote the new quantities related to photo-electric heating,
\begin{align}
    \widehat{\Phi}_i & \equiv \int_{\nu_i}^{\infty} \fheatE I_{\nu} e^{-\tau_{\nu}} \frac{d\nu}{h\nu} \\
    \widehat{\Psi}_i & \equiv \int_{\nu_i}^{\infty} \fheatE I_{\nu} e^{-\tau_{\nu}} d\nu .
\end{align}
Now, our ionization and heating coefficients read
\begin{align}
    \Gamma_i & = A_i \left[\Phi_i(\tau_{\nu}) - \Phi_i(\tau_{i,\nu}^{\prime}) \right] \label{eq:Gamma_PhiPsi_complex} \\
    \gamma_{ij} & = \frac{A_j}{h\nu_i} \left\{\widetilde{\Psi}_{ij}(\tau_{\nu}) - \widetilde{\Psi}_{ij}(\tau_{j,\nu}^{\prime}) - h \nu_j \left[\widetilde{\Phi}_{ij}(\tau_{\nu}) - \widetilde{\Phi}_{ij}(\tau_{j,\nu}^{\prime}) \right] \right\}  \label{eq:gamma_PhiPsi_complex} \\
    \Heat_i & = A_i \left\{\widehat{\Psi}_i(\tau_{\nu}) - \widehat{\Psi}_i(\tau_{i,\nu}^{\prime}) - h\nu_i \left[\widehat{\Phi}_i(\tau_{\nu}) - \widehat{\Phi}_i(\tau_{i,\nu}^{\prime}) \right] \right\} \label{eq:Heat_PhiPsi_complex} .  
\end{align}    
Various methods for handling secondary electrons are compared for a $10^5$ K blackbody source of $\dot{Q} = 5 \times 10^{48}$ ionizing photons $s^{-1}$ in Figure \ref{fig:SecondaryIonizationMethods}.

% Secondary electron methods
\begin{figure*}[htbp!]
\begin{center}
\includegraphics[width=0.98\textwidth]{figures/secondary_ionization_methods.eps}
\caption{Comparison of different methods for secondary ionization.  See \S\ref{sec:PhysicsParameters} for details on setting the secondary ionization method.}
\label{fig:SecondaryIonizationMethods}
\end{center}
\end{figure*}

% Cosmology
\subsection{Additional Complexity: Cosmological Expansion}
For an expanding universe, things are slightly more complicated, as ion densities will have an additional term $\propto (1 + z)^3$, and we must take into account Hubble cooling, and Compton scattering of CMB photons.

Let's first account for dilution of ion densities by expansion.  The number density of atoms at the cosmic mean density evolves as
\begin{align}
    \frac{dn}{dt} & = \frac{dn}{dz}\frac{dz}{dt} \nonumber \\
    & = \frac{d}{dz}\left(\bar{n}^0 (1 + z)^3 \right) \frac{dz}{dt} \nonumber \\
    & = 3 \bar{n}^0 (1 + z)^2 \frac{dz}{dt} .
\end{align}    
But, $dz/dt = H(z) (1 +z)$ and $\bar{n}^0 (1 + z)^3 = n$, so
\begin{equation}
    \frac{dn}{dt} = -3 n H(z) , \label{eq:dilution}
\end{equation}    
where we've inserted a minus sign to indicate that densities decrease with increasing time, or decreasing redshift.  The right-hand side of Equation \ref{eq:dilution} is added to Equations \ref{eq:HIIRateEquation}-\ref{eq:HeIIIRateEquation} when \texttt{CosmologicalExpansion = 1}.

We must also modify the minimum and maximum column densities for our integral tables.  To calculate what these should be for a simulation spanning the redshift interval $(z_i, z_f)$, minimum species fraction $x_{\mathrm{min}}$, and grid resolution (in proper coordinates) $dx$, do
\begin{align}
    N_{\HI}^{\mathrm{min}} & = \bar{n}_H^0 (1 + z_f)^3 x_{\mathrm{min}} dx   \\
    N_{\HI}^{\mathrm{max}} & = \bar{n}_H^0 (1 + z_i)^3 \Lbox
\end{align}

% Time-dependent source SEDs
\subsection{Additional Complexity: Time-Dependent Source SEDs}
Our lookup tables will inherit an additional dimension.  Goodie.

% Spectrum Normalization
\subsection{Spectrum Normalization}
For convenience, we generally work with the normalized SED $I_{\nu}$, such that 
\begin{equation}
    L_{\nu} = L_{\mathrm{bol}} I_{\nu} ,
\end{equation}    
and 
\begin{equation}
    \dot{Q}_{\nu} = L_{\nu} / h\nu .
\end{equation}    
However, in \S\ref{sec:Discretization}, we will discuss how \texttt{rt1d} can be used to verify choices of discrete SEDs -- when \texttt{DiscreteSpectrum = 1}, we must be careful about how we normalize a spectrum.

There are two primary ways we can go about normalizing a discrete spectrum consisting of $n_{\nu}$ frequency bins.

The first way is to manually supply the location of each bin (in eV) and the fraction of the bolometric luminosity allocated to that bin via the parameters \texttt{DiscreteSpectrumSED} and \texttt{DiscreteSpectrumRelLum}.  In \citet{Mirocha2012}, we denote these quantities as $h\nu_n$ and $I_n$, respectively, where $n = 1, 2,...n_{\nu}$, and do \textit{not} require the bolometric luminosity to remain conserved.  That is, we allow
\begin{equation}
    \sum_n I_n \neq 1.
\end{equation}    
This is because very high energy photons (depending on the spectrum) may not be absorbed at all over the column density regime or volume of interest.  It is important to note that even if $\sum_n I_n = 1$ exactly, the photon emissivity of a source constructed in this way will \textit{not} be conserved.

For example, when we simulate a $10^5$ K blackbody source emitting $5\times 10^{48}$ ionizing photons per second (as in RT06 problem \#2), the \texttt{RadiationSource} class figures out the bolometric luminosity (show equations here for reference), and assigns a fraction $I_n$ of that to bin $n$.  In doing this, 
\begin{equation}
    L_{\mathrm{bol}} \sum_n I_n / h\nu_n \neq 5 \times 10^{48} \ \mathrm{s}^{-1} .
\end{equation}

The second approach we can use is that used in previous work such as \citet{Aubert2008}, which requires us to define bandpasses, each defined by an average photon energy and effective bound-free absorption cross-section.  The details will be discussed in \S\ref{sec:Discretization}, but for now it is enough to know that the ionizing photon luminosity \textit{will} be preserved with this method, assuming that the bandpasses span the full emission interval of the radiation source of interest, and the bolometric luminosity will \textit{not} be preserved.  That is,
\begin{equation}
    \sum_n \dot{Q}_n h \nu_n \neq L_{\mathrm{bol}} .
\end{equation}

\subsection{Time-stepping Techniques}

\subsection{Parallelism}

% Lookup Tables
\subsection{Lookup Tables}
How does the dimensionality of our lookup tables depend on input parameters?

%%% 
%% Classes
%%%
\section{Classes}
In order of appearance in \texttt{RT1D.py}.

\subsection{\texttt{InitializeGrid}}
Initialize the temperature and ion densities for each cell in the grid.
\subsection{\texttt{InitializeIntegralTables}}
If \texttt{IntegralTabulation} = 1 (necessary when \texttt{DiscreteSpectrum} = 0), tabulates the quantities $\Phi_i$ and $\Psi_i$.  If \texttt{MultiSpecies} = 1, tabulation occurs for hydrogen and helium ions, and if \texttt{SecondaryIonization} $\ge 2$, also tabulate $\widetilde{\Phi}_i$, $\widetilde{\Psi}_i$, $\widehat{\Phi}_i$, and $\widehat{\Psi}_i$, adding an additional dimension for $x_{\mathrm{HII}}$.
\subsection{\texttt{RadiationSource}}
Initialize the radiation source.  Most importantly the \texttt{Spectrum} function, which returns the fraction of the bolometric luminosity emitted at energy $E$ (only argument).
\subsection{\texttt{Radiate}}
Calls \texttt{EvolvePhotons} routines, which loop over grid, call solver, and update ion densities and the temperature in each cell.
\subsection{\texttt{SolveRateEquations}}
ODE solver.
\subsection{\texttt{Interpolate}}
Performs linear, bi-linear, tri-linear, and eventually quad-linear interpolation on integral tables to extract ionization and heating coefficients as a function of column density.
\subsection{\texttt{WriteData}}
Write out data to HDF5.  Would be easy to add support for ASCII write-out.

%%% 
%% Parameters
%%%
\section{Parameters} \label{sec:Parameters}

\subsection{Problem Types}
These were added for convenience in January, 2011.  If you want to make changes to these test problems you can, just make sure \texttt{ProblemType} is the first (uncommented) line in your parameter file.  Additional parameters afterwards will be interpreted as usual.

\begin{description}

\item [\texttt{ProblemType = 0}] Single-zone.

\item [\texttt{ProblemType = 1}] This will set up a standard test with known analytic solution: the propagation of an ionization front in an isothermal, hydrogen-only, static universe.  The spectrum used is monochromatic at $E = 13.6 \ \text{eV}$, with photon luminosity $L_{\nu} = 5\times 10^{48} \ s^{-1}$.  The initial density, temperature, and box size are the same as the values used in Test 1 of Wise \& Abel 2010.

\item [\texttt{ProblemType = 2}] This problem is almost exactly the same as \texttt{ProblemType = 1}.  However, the temperature is allowed to evolve, and instead of using a monochromatic spectrum, we use a $10^5$ K blackbody spectrum.  Again, the initial density, temperature, box size, and energy groups are the same as the values used in Test 1 of Wise \& Abel 2010.

\item [\texttt{ProblemType = 2.1}] Same as \texttt{ProblemType = 2}, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Wise2011}.

\item [\texttt{ProblemType = 2.2}] Same as \texttt{ProblemType = 2}, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Mirocha2012}.

\item [\texttt{ProblemType = 3}] Test problem \# 3 from \cite{Iliev2006} -- I-front trapping in a dense clump and the formation of a shadow.  The setup is the same as in \texttt{ProblemType = 2}, except a 1D clump is initialized at $x = 0.76 L_{\mathrm{box}}$, overdensity $\delta = 200$, radius $0.12$kpc, and temperature $T = 40$K, and the radiation field is plane-parallel.

\item [\texttt{ProblemType = 3.1}] Same as \texttt{ProblemType = 3}, except we sample the continuous blackbody spectrum with the four energy groups used by \citet{Wise2011}.

\end{description}

\subsection{Initial Conditions}
\begin{description}

\item [\texttt{DensityProfile}]

\item [\texttt{TemperatureProfile}]

\item [\texttt{InitialHIIFraction}]

\end{description}

\subsection{Control Parameters}

\begin{description}

\item [\texttt{ODEatol}] Absolute tolerance of the ODE solver.  If ionized fractions are zero to this tolerance, they are reset to \texttt{MinimumSpeciesFraction}.  Combination of this parameter and \texttt{ODErtol} (below) determine whether or not the ODE time-step is cut in half. (Default = $10^{-5}$)

\item [\texttt{ODErtol}]
Relative tolerance of the ODE solver.  Combination of this parameter and \texttt{ODEatol} determine whether or not the ODE time-step is cut in half via the formulae $|y_2 - y_1| = \texttt{ODEatol} + \texttt{ODErtol}\times y_2$. (Default = $10^{-5}$)

\end{description}

\subsection{Source Parameters}

\begin{description}

\item \texttt{SourceType}
\begin{description}
    \item [0] Monochromatic source.
    \item [1] Star.  
    \item [2] PopIII star.  
    \item [3] Black hole.        
    \item [4] Absorbed power-law (column density given by \texttt{SpectrumAbsorbingColumn} parameter).
    \item [5] Multi-color disk + power-law. Relative normalization given by \texttt{SpectrumDiskFraction} parameter.                
\end{description}    
\end{description}

\subsection{Spectrum Parameters}
We keep these parameters separate from Source parameters to distinguish which parameters are allowed to vary with time, mass, luminosity, etc.  This separation was useful after renovating the \texttt{RadiationSource} class in May 2012, which made it so any parameter beginning with the word `Spectrum' can be a list, to allow construction of spectra consisting of an arbitrary number of components, each with uniquely specified parameters.

\subsection{Physics Parameters} \label{sec:PhysicsParameters}

\section{Analysis}
Since our datasets are small, we can read in the entire time evolution all at once.  To do this, type the following in a Python terminal:
\begin{verbatim}
import rt1d.analysis as rta
import pylab as pl

ds = rta.Analyze('./pf.dat')                # load parameter file and data
pl.loglog(ds.data[10].r, ds.data[10].x_HI)  # radius vs. neutral fraction
\end{verbatim}

%%%
%% Test Problems
%%%
\section{Test Problems}
In this section we focus on test problems 0-3 of the Radiative Transfer Comparison Project \citep[hereafter RT06;][]{Iliev2006}.

\subsection{Test \#0}
Single-zone ionization, heating, and cooling.

\begin{figure*}[htbp!]
\begin{center}
\includegraphics[width=0.98\textwidth]{figures/rt06_0.eps}
\caption{Single-zone ionization, heating, and cooling RT06 \#0.}
\label{fig:RT06_0}
\end{center}
\end{figure*}

\subsection{Test \#1}

\begin{figure*}[htbp!]
\begin{center}
\includegraphics[width=0.98\textwidth]{figures/rt06_1_convergence.eps}
\caption{Convergence test for RT06 \#1.}
\label{fig:RT06_1}
\end{center}
\end{figure*}

\subsection{Test \#2}

\begin{figure*}[htbp!]
\begin{center}
\includegraphics[width=0.98\textwidth]{figures/rt06_2.eps}
\caption{RT06 \#3.}
\label{fig:RT06_2}
\end{center}
\end{figure*}

\subsection{Test \#3}

\begin{figure*}[htbp!]
\begin{center}
\includegraphics[width=0.98\textwidth]{figures/rt06_3.eps}
\caption{RT06 \#3.}
\label{fig:RT06_3}
\end{center}
\end{figure*}

%%%
%% SED studies
%%%
\section{Verifying Discrete SEDs with \texttt{rt1d}} \label{sec:Discretization}
In \citet{Mirocha2012}, we used \texttt{rt1d} to verify new, optimally
selected discrete spectral energy distributions for sources of ultraviolet and
X-ray photons. As we optimize more and more SEDs for fully 3D cosmological
radiation-hydro simulations, \texttt{rt1d} remains a quick and easy tool for
verifying that the SEDs we're using are acceptable.

In the literature, discrete SEDs have been constructed using a `multi-group'
approach, which defines new absorption cross sections which represent
spectrum-weighted averages over broad bandpasses,
\begin{equation}
    \tilde{\sigma}_{i,n} \equiv \int_{\nu_n}^{\nu_{n+1}} \sigma_{i,\nu} \frac{I_{\nu}}{h\nu}  d\nu \left(\int_{\nu_n}^{\nu_{n+1}}\frac{I_{\nu}}{h\nu}d\nu \right)^{-1} ,
\end{equation}    
where the interval $(\nu_n, \nu_{n+1})$ represents the bandpass bounds for
discrete photon group $n$.

The emission energies are chosen in a similar way, 
\begin{equation}
    h\nu_n = \int_{h\nu_n}^{h\nu_{n+1}} I_{\nu} d\nu \left(\int_{h\nu_n}^{h\nu_{n+1}} \frac{I_{\nu}}{h\nu} d\nu \right)^{-1} . \label{eq:MeanPhotonEnergy}
\end{equation}    
It's worth keeping in mind that there are different ways to compute the mean
photon energy in each interval. Here, we've calculated by dividing a quantity
proportional to the bolometric energy output by the total number of ionizing
photons. We could also compute the expectation value, median value, or even
the peak of emission (e.g. for a blackbody). Equation
\ref{eq:MeanPhotonEnergy} is most often used in the literature, so we'll stick
to that.

%%%
%% Advanced usage
%%%
\section{Advanced Usage}
In \cite{Mirocha2012}

%%% 
%% Troubleshooting
%%%
\section{Troubleshooting}
If your results look horribly, horribly wrong, here are a few places where you could've gone wrong.  We'll start with things that may occur with the stable version of the code.

Now, if you modify or add anything in the source code (awesome by the way!), you're far more likely to encounter troubles.

\begin{description}
    \item[Totally Bogus Results] This is will almost certainly happen at some point in the development process. But, if it happens even after you're \textit{positive} the error was fixed, my guess is that you are using the integral tables that were generated with the previous, error-ridden version of the code.  Try removing the integral tables and re-running your simulation.
\end{description}

%%%
%% INTERACTIVE MODE
%%%
\section{Interactive Mode}
Is it worth doing this?  Would want it to look something like this:

\begin{verbatim}
import rt1d
import numpy as np
pf = rt1d.SetDefaultParameterValues()
r = np.linspace(3e19, 3e21, 100) # 100 grid cells between 0.01-1kpc. (optional)
data = rt1d.Shine(pf, r)    
\end{verbatim}

This would make it easy to:
\begin{itemize}
    \item Allow users to write their own rt1d `scripts.'
    \item Change source properties manually with time.
    \item Making arbitrarily structured grids.
    \item Very specific initial conditions.
\end{itemize}

What parameters must be set?
\begin{itemize}
    \item ProblemType
    \item GridDimensions
    \item MultiSpecies
    \item LengthUnits
    \item InitialDensity
    \item SourceParameters
\end{itemize}

\newpage
\bibliography{references}
\bibliographystyle{apj}

\end{document}